Estimating dark matter halo concentration using the integrated mass profile

### 2015-12-23 paper/paper.tex
nuevo texto en el paper
Update paper.tex
Un par de cambio menores en el texto (lineas 180 y 225)

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Halo concentration from the mass profile]{Estimating dark matter
  halo concentration using the integrated mass profile} 
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

The internal DM distribution in a halo can be parameterized through the density
profile.  
In a first approximation this profile is spherically symmetric and
its density only dependens on the radial coordinate. 
One the most popular radial parameterizations is  is the
Navarro-Frenk-White (NFW) profile 
\citep{NFW}.   
If one is not interested in the very central region of the
dark matter halo (where galaxy formation takes place, and where the
effects of baryon physics on dark matter distribution are still 
unknown) one can also assume that this profile is universal.
\citep{Navarro2010}.  
 This profile is a double power law in radius, where the transition
break happens at the so-called scale radius, $r_s$.  
The ratio between the scale radius and the halo virial radius $R_v$ is
known as the concentration $c=R_v/r_s$.

@@

The first method is straightforward to apply but presents two
disadvantages.  
First, it requires a large number of particles in
order to have a proper density estimate in each bin.  
This makes the method robust only for halos with at least $10^3$ particles.  
The second problem is that there is not a way to estimate the optimal
radial bin size, different choices produce different results for the
concentration.

@@

We explored halo shape and relaxation as a source of this
differences. We found shape does not bias the concentration
estimation; relaxation, quantified by the
offset between the center of mass and minimum of the potential, can
only explain at most $5\%$ of the offset between methods.  We note
that the highest concentration values correspond to the smallest halos
with $\sim 200$ particles, a regime where both the density and
velocity methods are reliable only the $5\%$ in a perfect spherical
halo.  From this we suggest that asserting with confidence the
systematic difference between the different methods should be
confirmed  a higher resolution simulation and larger statistics. We
leave this comparison for future work.

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Halo concentration from the mass profile]{Estimating dark matter
  halo concentration using the integrated mass profile} 
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{juan.munozc@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

The internal DM distribution in a halo can be parameterized through the density
profile.  
In a first approximation this profile is spherically symmetric and
its density only dependens on the radial coordinate. 
One the most popular radial parameterizations is the
Navarro-Frenk-White (NFW) profile 
\citep{NFW}.   
If one is not interested in the very central region of the
dark matter halo (where galaxy formation takes place, and where the
effects of baryon physics on dark matter distribution are still 
unknown) one can also assume that this profile is universal.
\citep{Navarro2010}.  
 This profile is a double power law in radius, where the transition
break happens at the so-called scale radius, $r_s$.  
The ratio between the scale radius and the halo virial radius $R_v$ is
known as the concentration $c=R_v/r_s$.

@@

The first method is straightforward to apply but presents two
disadvantages.  
First, it requires a large number of particles in
order to have a proper density estimate in each bin.  
This makes the method robust only for halos with at least $10^3$ particles.  
The second problem is that there is not a way to estimate the optimal
radial bin size, different choices may produce different results for the
concentration.

@@

We explored halo shape and relaxation as a source of this
differences. We found shape does not bias the concentration
estimation; relaxation, quantified by the
offset between the center of mass and minimum of the potential, can
only explain at most $5\%$ of the offset between methods.  We note
that the highest concentration values correspond to the smallest halos
with $\sim 200$ particles, a regime where both the density and
velocity methods are reliable only the $5\%$ in a perfect spherical
halo.  
This difference might actually signal a deviation from the NFW profile
due to the low resolution in the simulation. From this we suggest that
asserting with confidence the systematic difference between the
different methods should be confirmed  a higher resolution simulation
and larger statistics. We leave this comparison for future work.
